Selective Sequential Model Selection

### 2015-12-06 paper.tex
more editing, especially of simulation section; tweaked error rate tables code

\begin{table}[ht]
\centering
\input{tables/diabetes_edit.tex}
\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values. The bold annotation in each column marks the model selected by ForwardStop with $\alpha=0.1$. For the max-$t$ $p$-values, this stopping rule gives exact FDR control at the 10\% level.}
\label{tab:diab}
\end{table}

@@

\subsection{Constraints}\label{sec:constraints}

@@

A similar result also holds for the class of $\ell_1$-regularized exponential family models. % including the graphical LASSO.

@@

For each step we compute saturated-model $p$-values, max-$t$ and max-$z$ $p$-values, and nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable. \WFcomment{Do we really need to condition on the signs? There's just two polytopes in each case, right?} Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first eight steps, conditional on the event that the null hypothesis is false. It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values, especially for the first five steps, where they have near-perfect power. The nominal $p$-values are also quite powerful, but they do not have the correct level.

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$. The max-$t$ approaches are much more powerful than the saturated-model test. The nominal test appears to be powerful, 
  but is not $U[0,1]$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
  \label{fig:simulation_null_false}
\end{figure}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.  The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
  \label{fig:simulation_null_true}
\end{figure}

 \RJTcomment{The lines in Fig 6 are super thin, thinner than Fig 5?}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
  \caption{\em CDFs of nominal (black), max-$z$ (red), max-$t$ $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model. Here, none of the methods produce uniform $p$-values. The null hypothesis is false in most cases and so --- in our model-centric point of view --- rejection is the desired outcome.}
  \label{fig:simulation_noise_var}
\end{figure}

@@

If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method. Table~\ref{tab:stopping05} reports several model-selection metrics for each combination of test and stopping rule carried out at level $\alpha=0.05$, while Table~\ref{tab:stopping20} reports the same results for the more liberal $\alpha=0.2$. The metrics are $p_{\text{screen}}$, the probability of selecting a correct model; the (model-wise) FWER, both marginally and conditional on screening, the model-wise FDR and variable-wise $\text{FDR}^{\text{full}}$, and the average number of signal variables selected. Note that $\text{FDR}^{\text{full}}$ is not explicitly controlled by any of the methods other than knockoff+, but we might nevertheless hope to perform reasonably well. Also, note that the model-wise FWER and FDR and not defined for knockoffs, since it does not select variables sequentially, and never formally tests any model.

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Fri May  1 09:29:28 2015
%\begin{table}[ht]
 % \centering
 % \begin{tabular}{llcccc}
  %  \hline
  %  Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
 %   & $\text{FDR}_{\text{mod}}$ 
  %  & $\text{FDR}_{\text{var}}$ \\ 
 %   \hline
    %Selected & Basic & .290 & .000 & .002 & .039 \\ 
   % Selected & Forward & .559 & .027 & .020 & .066 \\ 
   % Selected & Strong & .041 & .014 & .008 & .042 \\ 
  %  \hline
   % Saturated & Basic & .000 & .000 & .000 & .028 \\ 
  %  Saturated & Forward & .014 & .000 & .000 & .030 \\ 
  %  Saturated & Strong & .000 & .000 & .000 & .032 \\ 
  %  \hline
  %  Knockoffs & & .000 & --- & --- & .231 \\ 
  %  \hline
 % \end{tabular}
 % \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
%\end{table}

\begin{table}[ht]
\centering
\input{tables/error_rates_05.tex}
\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}. Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
\label{tab:stopping05}
\end{table}

\begin{table}[ht]
\centering
\input{tables/error_rates_20.tex}
\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}. Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
\label{tab:stopping20}
\end{table}

We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism. Although not guaranteed by the theory, the $p$-values from the saturated model do control model-wise FDR in this example.  The max-$t$ rules show higher probabilities of selecting the true model, especially using ForwardStop.  Lastly, the knockoff method does not yield variable-wise FDR control. While the more conservative knockoff+ method does achieve variable-wise FDR control, this comes at a cost of very low power. Because of the discrete nature of knockoff and knockoff+, these two methods might perform better in an example with a larger number of signal variables to find.

@@

Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
\begin{figure}[htp]
\centering
  \includegraphics[width=\textwidth]{figs/comparison.pdf}
  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.  Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
}
\label{fig:comparison}
\end{figure}
We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples. As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate. Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low for accept/reject to be practical.

>>>>

\begin{table}[ht]
\centering
\input{tables/diabetes_edit.tex}
\caption[tab:diab]{Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values. The bold annotation in each column marks the model selected by ForwardStop with $\alpha=0.1$. For the max-$t$ $p$-values, this stopping rule gives exact FDR control at the 10\% level.}
\label{tab:diab}
\end{table}

@@

\subsection{Constraints for the Greedy Likelihood and Next-Entry Tests}\label{sec:constraints}

@@

A similar result also holds for the class of $\ell_1$-regularized exponential family models.

@@

For each step we compute saturated-model $p$-values, max-$z$ $p$-values, and nominal (unadjusted) $p$-values, conditioning on the signs of the active variables as proposed by \citet{lee2013exact}. Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first eight steps, conditional on the event that the null hypothesis is false. It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values, especially for the first five steps, where they have near-perfect power. The nominal $p$-values are also quite powerful, but they do not have the correct level.

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
  \caption{  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$. The max-$t$ approaches are much more powerful than the saturated-model test. The nominal test appears to be powerful, 
  but is not $U[0,1]$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
  \label{fig:simulation_null_false}
\end{figure}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
  \caption{  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.  The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
  \label{fig:simulation_null_true}
\end{figure}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
  \caption{ CDFs of nominal (black), max-$z$ (red), max-$t$ $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model. Here, none of the methods produce uniform $p$-values. The null hypothesis is false in most cases and so --- in our model-centric point of view --- rejection is the desired outcome.}
  \label{fig:simulation_noise_var}
\end{figure}

@@

If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method. Table~\ref{tab:stopping05} reports several model-selection metrics for each combination of test and stopping rule carried out at level $\alpha=0.05$, while Table~\ref{tab:stopping20} reports the same results for the more liberal $\alpha=0.2$. The max-$t$ test, which we could use in place of the max-$z$ test if $\sigma^2$ were unknown, is included for comparison. The max-$z$-identify test was also included in the simulation, but not shown in Tables~\ref{tab:stopping05} and \ref{tab:stopping20} because its performance on all metrics was virtually identical to the max-$z$ test.

The metrics are $\P(\hk \geq k_0)$, the probability of selecting a correct model; the (model-wise) FWER; the FWER conditional on $\hk \geq k_0$, labeled cFWER; the model-wise FDR and variable-wise $\text{FDR}^{\text{full}}$; and the average number of full-model signal variables selected, labeled $\E[S^{\text{full}}]$ (where $S^{\text{full}} = R-V^{\text{full}}$). Note that $\text{FDR}^{\text{full}}$ is not explicitly controlled by any of the methods other than knockoff+, but we might nevertheless hope to perform reasonably well. Also, note that the model-wise FWER and FDR and not defined for knockoffs, since it does not select variables sequentially, and never formally tests any model.

\begin{table}[ht]
\centering
\input{tables/error_rates_05.tex}
\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  at level $\alpha=0.05$. Values theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type. The largest standard error in each of the six columns is 0.01, 0.01, 0.02, 0.004, 0.003, and 0.05.}
\label{tab:stopping05}
\end{table}

\begin{table}[ht]
\centering
\input{tables/error_rates_20.tex}
\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals, at level $\alpha=0.2$. Values theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type. The largest standard error in each of the six columns is 0.01, 0.01, 0.03, 0.004, 0.004, and 0.07.}
\label{tab:stopping20}
\end{table}

We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$z$ and max-$t$, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism. Although not guaranteed by the theory, the $p$-values from the saturated model do control model-wise FDR in this example.  The max-$z$ method is more powerful, showing higher probabilities of selecting a correct model, especially when using ForwardStop. The max-$t$ method is also powerful but slightly less so than the max-$z$, owing to the additional nuisance variable $\sigma^2$.

The knockoff method does not yield variable-wise FDR control. While the more conservative knockoff+ method does achieve variable-wise FDR control, this comes at a cost of very low power. Because of the discrete nature of knockoff and knockoff+, these two methods might perform better in an example with a larger number of signal variables to find.

@@

Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
\begin{figure}[htp]
\centering
  \includegraphics[width=\textwidth]{figs/comparison.pdf}
  \caption{ Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.  Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
}
\label{fig:comparison}
\end{figure}
We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples. As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate. Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low for accept/reject to be practical.
